Compiler and Runtime Support for Functions-as-a-Service (FaaS)

### 2021-12-09 paper/paper.tex
[paper]: add more eval text, figures and tables, add conclusion

Finally, all workloads are implemented in \texttt{C} and only use \textit{libc} \texttt{malloc} and 
\texttt{free}. Although most FaaS workloads are implemented in higher-level languages, the 
simplicity of the dynamic memory interface makes it very simple for the prototype compiler and 
runtime to target, as all \ALLS\ and \DALLS\ are obviously explicit. Targeting higher-level 
languages remains a future work.

@@

We assess the compiler transforms for the custom allocator both statically and dynamically. Figure X
shows the number of eligible static \ALLS\ in the baseline IR for each kernel and the number of \ALLS\
that were transformed to invoke the allocator.

Next, we evaluate the performance of the transformed binaries against the baseline binaries (i.e.
without transformation, compiled straight from source code) for each kernel. Each kernel was 
run 10 times via \texttt{perf stat} and the cycle counts were recorded. Figure X shows the comparative
overhead ratio (in terms of cycles).

Finally, we break down the execution times for each major piece of the allocator runtime and
highlight worst-case observed times (also in cycles). Figure X shows this for each essential
method of the runtime.

\begin{figure}
        \small
        \centering
        \begin{tabular}{ | m{5cm} | m{9cm} | m{3cm} | } 
          \hline
            \texttt{Init} & Initialize the global allocator (that manages bump allocators) and compiler directed pool & 20418 \\ 
          \hline
            \texttt{AddAllocator} & Initialize a new bump allocator and add it to the global allocator & 27170 \\ 
          \hline
            \texttt{AllocateFromPool} & Fetching a pointer from the compiler-directed memory pool (wrapped as a function) & 52 \\ 
          \hline
            \texttt{AllocateFromBump} & Fetching a pointer from a bump allocator & 1372 \\ 
          \hline
            \texttt{AllocateFromBumpWithRuntimeInit} & Fetching a pointer from a bump allocator with allocator initialization, if necessary & 1372 \\ 
          \hline
        \end{tabular}
        \caption{Worst-case cycle times for essential methods in allocator runtime.}
        \label{fig:apps}
\end{figure}

@@

\section{Future Work}

\section{Conclusion}

>>>>

Finally, all workloads are implemented in \texttt{C} and only use \textit{libc} \texttt{malloc} and 
\texttt{free}. Although most FaaS workloads are implemented in higher-level languages, the 
simplicity of the dynamic memory interface makes it very simple for the prototype compiler and 
runtime to target, as all \ALLS\ and \DALLS\ are obviously explicit. Targeting higher-level 
languages remains a future work. All inputs are placed on the stack as a simulation of the 
JIT/binary translation-based solution for embedding inputs into the binary before a FaaS 
instance executes.

@@

We assess the compiler transforms for the custom allocator both statically and dynamically. Figure~\ref{fig:static}
shows the number of eligible static \ALLS\ (\texttt{malloc}s column) in the baseline IR for each kernel and 
the number of \ALLS\ that were transformed to invoke the allocator ("Transformed" column). For matrix 
multiplication, the compiler aggressively unrolled the entire allocation loop for the result matrix
and inlined the entire \texttt{multiply} function. Consequently, all \SIZES\ were statically analyzable
and all \ALLS\ were transformed to fetch from the compiler-directed pool. For breadth first traversal and Huffman 
coding, almost all \ALLS\ were transformed to allocate from the bump allocator because most \ALLS\
were found in loops but had statically analyzable \SIZES . The \ALLS\ not transformed in
the Huffman coding kernel were all in functions that were inlined. The inlined static \ALLS\ were
all transformed.

\begin{figure}[htp]
    \centering
    \begin{minipage}{0.45\textwidth}
        \small
        \centering
        \begin{tabular}{ | m{5cm} | m{1cm} | m{2cm} | } 
          \hline
            Kernel & \texttt{malloc}s & Transformed \\
          \hline\hline
            Matrix Multiplication & 60 & 60 \\ 
          \hline
            Compressed-Row-Storage (CSR) & 3 & 0 \\ 
          \hline
            Breadth First Traversal & 7 & 7  \\ 
          \hline
            Huffman Coding & 17 & 10 \\ 
          \hline
        \end{tabular}
        \caption{Kernels used for evaluation.}
        \label{fig:static}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.88\textwidth]{figs/graph.png} 
        \caption{Overhead ratio for transformed kernels (with allocator runtime)
        vs. baseline kernels (no transform). Ratio compares cycle counts.}  
	    \label{fig:perf}
    \end{minipage}\hfill
\end{figure}

Next, we evaluate the performance of the transformed binaries against the baseline binaries (i.e.
without transformation, compiled straight from source code) for each kernel. Each kernel was 
run 10 times via \texttt{perf stat} and the cycle counts were recorded. Figure~\ref{perf} shows the 
comparative overhead ratio (in terms of cycles). Here, all kernels have minimal overhead; although
matrix multiplication, in particular, should be faster. The lack of speedups could be attributed
to a very low-latency \texttt{malloc} coupled with extra initialization and runtime overhead. This
is explored in Figure~\ref{cycles}. We break down the execution times for each major piece of the 
allocator runtime and highlight worst-case observed times (also in cycles). These overheads are
fairly substantial for the work that each method is performing. Some of the overhead can be due
to extra function calls because wrapper methods need to be exposed to the compiler for streamlined
instrumentation. Additionally, the data structures (\texttt{std::map}) used in the runtime and 
initialization costs may drive up overheads. Many of these overheads could be optimized by 
careful engineering.

\begin{figure} [htp]
        \small
        \centering
        \begin{tabular}{ | m{5cm} | m{9cm} | m{1cm} | } 
          \hline
            \texttt{Init} & Initialize the global allocator (that manages bump allocators) and compiler directed pool & 20418 \\ 
          \hline
            \texttt{AddAllocator} & Initialize a new bump allocator and add it to the global allocator & 27170 \\ 
          \hline
            \texttt{AllocateFromPool} & Fetching a pointer from the compiler-directed memory pool (wrapped as a function) & 52 \\ 
          \hline
            \texttt{AllocateFromBump} & Fetching a pointer from a bump allocator & 1372 \\ 
          \hline
            \texttt{AllocateFromBumpWithRuntimeInit} & Fetching a pointer from a bump allocator with allocator initialization, if necessary & 1372 \\ 
          \hline
        \end{tabular}
        \caption{Worst-case cycle times for essential methods in allocator runtime.}
        \label{fig:cycles}
\end{figure}

@@

\section{Conclusion}
This paper shows a \crcd\ that analyzes and optimizes for dynamic memory usage in 
small FaaS workloads. The system provides a profiler, static allocation tracking and 
analysis, and a custom memory allocator controlled by both the compiler and runtime
tooptimize for dynamic memory. Across a small set of kernels, the compiler is 
successful at analyzing and transforming allocation candidates to use the custom
allocator, and the system achieves comparable performance compared to the baseline 
applications.
